---
title: "p8105_hw6_jl6047"
author: "Jiayi Luo"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(purrr)
library(modelr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "right"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```
## Problem 2 

```{r, warning=FALSE, message=FALSE}
# Import the dataset

homicide_raw <- read_csv(url("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"), na = c("", "NA", "Unknown"))

# Clean dataset,create variables city_state and resolved
# resolved: 0 unresolved 1 resolved
# Make sure victim age is numeric
# Omit cities without victim race reports and error city
# Limit victim_race to white or black
homicide_df = homicide_raw %>% 
  janitor::clean_names() %>%
  mutate(reported_date = as.Date(as.character(reported_date), format = "%Y%m%d"),
         victim_age = as.numeric(victim_age),
         victim_race = fct_relevel(victim_race, "White"),
         city_state = str_c(city, state, sep = ", "),
         resolved = case_when(
           disposition == "Closed without arrest" ~ 0,
           disposition == "Open/No arrest" ~ 0,
           disposition == "Closed by arrest" ~ 1)) %>% 
        relocate(city_state) %>% 
        filter(!city_state %in% c("Dallas, TX","Phoenix, AZ","Kansas City, MO","Tulsa, AL"),
         victim_race %in% c("White","Black"))
```


```{r, warning=FALSE, message=FALSE}
# Use  foir baltimore with resolved vs unresolved
# Save the output to R project

baltimore_logistic = homicide_df %>%
    filter(city_state == "Baltimore, MD") %>% 
    glm(resolved ~ victim_age + victim_race + victim_sex, data = ., family = binomial(link = "logit")) 

# Broom tidy
# Obtain estimates of OR for solving homicides comparing males vs female

baltimore_logistic %>% 
  broom::tidy(conf.int = T) %>% 
  mutate(OR = exp(estimate),
         CI_lower = exp(exp(conf.low)),
         CI_upper = exp(exp(conf.high)),
         p_val = rstatix::p_format(p.value, digits = 2)) %>% 
  select(term, OR, CI_lower,CI_upper, p_val) %>% 
  mutate(term = str_replace(term, "victim_age", "Age"),
         term = str_replace(term, "victim_race", "Race: "),
         term = str_replace(term, "victim_sex", "Sex: ")) %>% 
  knitr::kable(digits = 3, align = "lccc", 
               col.names = c("Term", "Estimated adjusted OR", "CI low", "CI high", "p-value"))
```

* In Baltimore, MD, controlling for all other variables, the homicides are significantly less like to be resolved in males compared to females.

```{r, warning=FALSE, message=FALSE}
#  Extract the adjusted odds ratio and CI 
#  Compare male victims with female victims from all cities this time.
allcities_logistic = homicide_df %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = map(data, ~glm(resolved ~ victim_age + victim_race + victim_sex, data = ., family = binomial(link = "logit"))),
    results = map(models, ~broom::tidy(.x, conf.int = T))) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(conf.low),
    CI_upper = exp(conf.high),
    p_val = rstatix::p_format(p.value, digits = 2)
  ) %>% 
  filter(term == "victim_sexMale") %>% 
  select(city_state, OR, CI_lower,CI_upper, p_val) 
allcities_logistic %>% 
  knitr::kable(digits = 3, align = "llccc", col.names = c("City", "Estimated adjusted OR", "CI low", "CI high", "p-value"))
```

```{r, warning=FALSE, message=FALSE}
# Create the plot showing the estimated ORs and CIs for each city
allcities_logistic %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>%
  ggplot(aes(x = city_state, y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(x = "City", y = "Estimated OR with CI")
```


## Problem 3
```{r, message=FALSE}
birthweight = read_csv("./birthweight.csv")
```

### Data Preprocessing

```{r}
# Load data
# clean data
# Create categorical variables for babysex,malform, frace and mrace
birthweight_tidy =
  birthweight %>% 
  janitor::clean_names() %>%
  mutate(across(.cols = c(babysex, frace, malform, mrace), as.factor)) %>%
  mutate(babysex = case_when(babysex == 1 ~"male", babysex == 2 ~"female"),
         malform = case_when(malform == 0 ~"absent", malform == 1 ~ "present"),
         frace = recode(frace, 
                        "1" = "White", 
                        "2" = "Black", 
                        "3" = "Asian", 
                        "4" = "Puerto Rican", 
                        "8" = "Other", 
                        "9" = "Unknown"),
         mrace = recode(mrace, 
                        "1" = "White", 
                        "2" = "Black", 
                        "3" = "Asian", 
                        "4" = "Puerto Rican", 
                        "8" = "Other"))
# check missing values
sum(is.na(birthweight_tidy))
```


```{r, message=FALSE}
# Propose a Regression Model for Birthweight
# Select variables with stepwise method
mult_fit = lm(bwt ~ ., data=birthweight_tidy)
step(mult_fit, direction="both")
```

After stepwise selection,
variables `babysex`, `bhead`, `blength`, `delwt`, `fincome`, `gaweeks`, `mheight`, `mrace`, `parity`, `ppwt` and `smoken` are left.

```{r}
#run linear regression to check fitness of the model
mult_fit_1 = lm(bwt ~  babysex + bhead + blength + delwt + fincome + gaweeks + mheight + mrace + parity + ppwt + smoken,
              data=birthweight_tidy)
summary(mult_fit_1)
```


```{r, message=FALSE, warning=FALSE}
fit_var = birthweight_tidy %>%
  select(bhead, blength, delwt, fincome, gaweeks, mheight, parity, ppwt, smoken)
# Check correlation matrix 
cor(fit_var)
# Create a scatter plot matrix
pairs(fit_var)
```

Conclusion:
There could exist collinearity between delwt and ppwt, also bhead and blength.
So I decide to drop one of each from the collinear groups, `ppwt` and `blength` whose p-values are relatively larger, so drop these 2 to create a new model.

```{r}
#model 2
mult_fit_2 = lm(bwt ~  babysex + bhead + delwt + fincome + gaweeks + mheight + mrace + parity + smoken,
              data=birthweight_tidy)
summary(mult_fit_2)
```

Next, I will show a plot of model residuals against fitted values.

```{r, message=FALSE}
birthweight_tidy %>%
  add_residuals(mult_fit_2) %>%
  add_predictions(mult_fit_2) %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Model Residuals vs. Fitted Values",
       x = "Fitted Values", y = "Model Residuals") 
```


The constant variance assumption is generally satisfied with most residuals scattered around 0 and extremes appear when prediciton is below 1000, of which are a small portion.

```{r,message=FALSE, warning=FALSE}
# Compare the models
fit_com_1 = lm(bwt ~ blength + gaweeks, data = birthweight_tidy)
fit_com_2 = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex,
               data = birthweight_tidy)
# Cross Validation 
cv_df = 
  crossv_mc(birthweight_tidy, 100) %>% 
    mutate(
        train = map(train, as.tibble),
        test = map(test,as.tibble)
    )  %>%
  mutate(
    model_fit1  = map(train, ~lm(bwt ~ babysex + bhead + delwt + fincome + gaweeks + mheight + mrace + parity + smoken,
                                 data = birthweight_tidy)),
    model_fit2  = map(train, ~lm(bwt ~ blength + gaweeks, data = birthweight_tidy)),
    model_fit3  = map(train, ~lm(bwt ~ blength + bhead + babysex + blength*bhead + blength* babysex + bhead*babysex + blength*bhead*babysex, data = birthweight_tidy))) %>% 
  mutate(
    rmse_1 = map2_dbl(model_fit1, test, ~rmse(model = .x, data = .y)),
    rmse_2 = map2_dbl(model_fit2 , test, ~rmse(model = .x, data = .y)),
    rmse_3 = map2_dbl(model_fit3, test, ~rmse(model = .x, data = .y))) 
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin() +
  labs(title = "Cross-Validated Prediction Error Comparison Graph", 
       x = "Models", 
       y = "RMSE")  +
  scale_x_discrete(labels = c("My Model", "Test Model 1", "Test Model 2")) 
```

With the lowest RMSE, test model 1 is potentially the most fitted model among these three models and can be further explored.